---
title: "Assignment, Week 4: District crime"
author: "Colleen Cronin"
date: "September 14, 2023"
output: html_document
---

<!-- Please knit this right away! The additional information for (html and css)
are provided to help you build on your current R Markdown toolbox. You will not
be responsible for learning these additional tools at this point. -->

<!-- The head tag, <head>, is a container for metadata in an html document. 
We can use it to define styles and do lots of other cool things things. -->
<head>
<!-- The link tag, <link>, creates a relationship between this file and
an external source. Here, I'm linking font-awesome so I can include a 
couple of icons that I like (user-secret and user-circle) -->
<link 
  rel="stylesheet" 
  href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">
<!-- Here I'm linking to a javascript file that will give us access to font awesome -->
  <script src="https://kit.fontawesome.com/03064bdd6c.js" crossorigin="anonymous"></script>
</head>
<!-- Note in the above that some html tags, like <head> need to be closed 
with </head>. Others, like <link> don't need a closing tag. -->

```{css, echo = FALSE}
.score {
  font-weight: bold;
  font-size: 16px;
  color: red;
  background-color: #ffff80;
}

.subscore {
  font-size: 14px;
  color: red;
}

/* Accordion style, the accordion is the section "Functions that you may use in this assignment" */
  
.accordion {
  background-color: #e6f0ff;
  color: #000000;
  cursor: pointer;
  padding: 18px;
  width: 100%;
  border-style: solid;
  border-width: 2px;
  border-color: #ffffff;
  text-align: left;
  outline: none;
  font-size: 15px;
  transition: 0.4s;
}

/* When you hover your mouse over the accordion button, it changes color. */

.active, .accordion:hover {
  background-color: #80b3ff; 
  color: #ffffff;
  font-weight: bold;
}

/* This is where the content of the accordion is placed. */

.panel {
  padding: 0 18px;
  display: none;
  background-color: #eeffff;
  overflow: hidden;
}

/* Here, I add a custom class for area surrounding the graded portions of this
assignment.*/

.now_you {
  border-left: 8px solid #292093;
  background-color: #e6ffff;
  padding: 16px;
  margin-top: 15px;
  margin-bottom: 15px;
}

/* This is where tips and tricks are formatted. */

.mysecret {
  border: 1px solid #999999;
  border-style: solid; 
  background-color: #efefef;
  padding: 20px; 
  border-radius: 25px;
  margin-top: 15px;
  margin-bottom: 15px;
}

h1, h2 {
  margin-bottom: 4px;
}

hr {
  margin-top: 0px;
}
```

<!-- Here is the setup for knitr. I had to include tidyverse so I could
include a tidyverse example with my Important warning statement below-->

```{r knitr_setup, include = FALSE}
options(knitr.table.format = "html")

knitr::opts_chunk$set(error = TRUE)
```

<hr>

## Overview
<hr>

<a title = "The city">
  <img  src = "https://dcist.com/wp-content/uploads/sites/3/2019/10/jeffersonmemorial-768x511.jpg"
        alt="Jefferson memorial"
        style = "float: right; 
                 padding-left: 1em; 
                 padding-bottom: 0.5em; 
                 padding-right: 0em; 
                 width: 40%; 
                 height: auto;">
  </img>
</a>

In this exercise, you will explore crime data from the District of Columbia (a.k.a., Washington, DC) from October 2019 to October 2021. Each crime record includes the geographic coordinates of the crime (recorded in EPSG 4326), the date and time it occurred, whether it was a *violent* or *property* crime, and the type of offense. Your job will be to generate maps of these crimes across the District.

This assignment will give you an opportunity to test your *sf*, *purrr::map*, and *tmap* skills!

## Grading
<hr>

The points allotted for each question are provided in highlighted red bold text (e.g., <span class = "score">[1.0]</span>) within the question itself. When applicable, total points for a question may represent the sum of individually graded components, which are provided in red text (e.g., <span class = "subscore">[1.0]</span>).

Points may be deducted from each question's total: 

* [[50% per violation]]{class="subscore"} Use only functions listed in "Functions that you may use in this assignment";
* [[50%]]{class="subscore"} Include only assignments specified in the question;
* [[10%]]{class="subscore"} Ensure that all code is properly indented;
* [[10%]]{class="subscore"} Ensure that your code follows modern coding conventions (see *Best practices in modern R coding*);
* [[10%]]{class="subscore"} Ensure that all code chunks are formatted as `eval = TRUE` (but see below);
* [[10%]]{class="subscore"} Ensure that the class of your response is consistent with the wording of the question (e.g., if you are asked to return a "value" or "vector", ensure that the resultant object is not a data frame).

*Note: The maximum deduction is the total points value for a given question*

:::{class="mysecret"}
<i class="fas fa-user-secret fa-2x"></i> Pay careful attention to the format of your code -- not following the rules above will cost you lots of little points (and some big ones) that can really add up!
:::

In addition to points allotted per question, you must ensure that your R Markdown document **runs out-of-the-box** [[25% off of the total grade]]{class="subscore"} -- in other words, the document will knit without error. Some tips for doing so:

* Do not maintain your project folder in a location that is backed up by an online directory (e.g., Dropbox, iCloud).
* Ensure that all file paths are equivalent to those in this document (e.g., in `source()` or `read_csv()`).
* Do not use `setwd()` in your code (***Never** use `setwd()`!*)
* If one of your code chunks contains an error, include `eval = FALSE` in the options for that chunk. Otherwise, ensure all of your code chunks options are `eval = TRUE`.
* Prior to submission, clear all objects from your global environment and try to knit your document again.

:::{class="mysecret"}
<i class="fas fa-user-secret fa-2x"></i> Click the blue button below to view the functions that you may use in completing this problem set. Make sure that you know what each function does (use `?[function name]` if you do not). Do not use any functions outside of this list!
:::

<div>
<button class="accordion">Functions that you may use in this assignment</button>
<div class="panel">

In this assignment, you may use **only** the following R functions (*Note: If you are unclear on what a given function does, use `?` to view the help file!*):

* base::`()`
* base::`=`
* base::`<-`
* base::`==`
* base::`!`
* base::`~`
* base::`$`
* base::c
* base::library
* base::list
* dplyr::left_join
* dplyr::n
* dplyr::select
* dplyr::summarize
* lubridate::year
* magrittr::`%>%`
* purrr::map
* readr::read_csv
* tidyr::pivot_wider
* sf::st_as_sf
* sf::st_crs
* sf::st_join
* sf::st_read
* sf::st_transform
* tibble::as_tibble
* tmap::`+`
* tmap::tmap_mode
* tmap::tm_basemap
* tmap::tm_dots
* tmap::tm_polygons
* tmap::tm_shape

<span class = "myhint">**Note**: The packages dplyr, ggplot2, lubridate, magrittr, purrr, readr, tidyr, and tibble are all part of the tidyverse metapackage and are loaded with `library(tidyverse)`.</span>

</div>
</div>

<hr>

## Getting started
<hr>

<div class = "now_you">

1\. [[0.5]]{class="score"} Save and knit this document:

* [[0.1]]{class="subscore"} Replace my name in the YAML header with yours
* [[0.1]]{class="subscore"} Add the current date in the YAML header
* [[0.3]]{class="subscore"} Save the *.rmd* file in the **output** folder of your project as (*but replace my name with yours*): `problem_set_4_Evans_Brian.rmd`

2\. [[0.5]]{class="score"} Set up your session:

* [[0.25]]{class="subscore"} Load the `sf`, `tmap`, and `tidyverse` libraries;

```{r question_2_libraries, results = 'hide', message = FALSE, warning = FALSE}

library(sf)
library(tmap)
library(tidyverse)

```

* [[0.25]]{class="subscore"} Set the tmap mode to "plot".

```{r question_2_tmap_mode, results = 'hide', message = FALSE, warning = FALSE}

tmap_mode('plot')

```

</div>

## Read and pre-process data
<hr>

We will use the shapefile dataset `dc_census.geojson` as a template upon which we will display crimes in the District of Columbia. The coordinate reference system (CRS) of these data is EPSG 32618 (Universal Transverse Mercator, zone 18N). The field `geoid` is the primary key for each polygon in the dataset and is the only field that we will use for this problem set.

<div class = "now_you">

3\.	[[0.75]]{class="score"}  As parsimoniously as possible, read in the census data (`dc_census.geojson`) [[0.15]]{class="subscore"}, and:

* [[0.25]]{class="subscore"} Remove all fields except `GEOID`;
* [[0.25]]{class="subscore"} Change the field name `GEOID` from upper to lower case;
* [[0.1]]{class="subscore"} Assign to your global environment with the name `census`.

```{r question_3, results = 'hide'}

census <-
st_read('data/raw/shapefiles/dc_census.geojson') %>% 
  select(geoid = GEOID)

```
</div>

Crime data, `dc_crimes.csv`, were obtained from <a href="https://opendata.dc.gov/" target="_blank">Open Data DC</a>. This is a tabular dataset and each row represents the record for an individual crime committed. Fields (columns) include:

* `id`: The primary key for each observation;
* `longitude` and `latitude`: The geographic coordinates of crimes, recorded using a handheld GPS unit set to EPSG 4326 (Geodetic CRS, World Geodetic System 1984);
* `date_time`: The date and time a crime occurred using the format yyyy-mm-dd hh:mm:ss (International Organization for Standardization 8601);
* `offense`: The type of crime committed;
* `offense_group`: Categories for offenses -- "property" or "violent" crimes.

Our goal in this problem set is to evaluate the number and spatial distribution of violent crimes that were committed in 2020.

<div class = "now_you">
4\.	[[1.25]]{class="score"} Read in the crimes dataset (`dc_crimes.csv`) [[0.15]]{class="subscore"}, and:

* [[0.25]]{class="subscore"} Subset crimes to where the `offense_group` is categorized as "violent" and the crime was committed in 2020;
* [[0.25]]{class="subscore"} Remove the fields `offense_group` and `date_time`;
* [[0.5]]{class="subscore"} Convert to an `sf` object with the same CRS as `census`;
* [[0.1]]{class="subscore"} Assign to your global environment with the name `crimes`.

```{r question_4}

crimes <-
  read_csv('data/raw/dc_crimes.csv') %>%
  filter(offense_group == "violent",
         lubridate::year(date_time) == 2020) %>%
  select(!offense_group,!date_time) %>%
  st_as_sf(coords = c('longitude', 'latitude'),
           crs = 4326) %>%
  st_transform(st_crs(census))

```
</div>

## Summarize crimes by census tract
<hr>

Later this problem set (questions 8 and 9), you will plot number and spatial distribution of crimes by census tract and type of offense. To do so, we must tabulate the number of crimes for each tract.

<div class = "now_you">
5\. [[2.0]]{class="score"} Create a shapefile that describes the number of crimes per census tract and offense:

a. [[0.5]]{class="subscore"} Join the `census` data to the `crimes` dataset;
b. [[0.2]]{class="subscore"} Convert the resultant object to a tibble;
c. [[0.5]]{class="subscore"} Calculate the number of crimes per `geoid` and `offense`;
d. [[0.5]]{class="subscore"} Reshape the resultant object such that the rows represent census tracts and column names are: `geoid`, `robbery`, `assault w/dangerous weapon`, `sex abuse`, and `homicide`;
e. [[0.2]]{class="subscore"} Convert the resultant object to a shapefile;
f. [[0.1]]{class="subscore"} Globally assign the resultant object as `crimes_by_census_tract`.

```{r census_crimes}

crimes_by_census_tract <-
  census %>%
  st_join(crimes,
          by = c("geoid" = "id")) %>%
  as_tibble() %>%
  summarize(n = n(),
            .by = c("geoid", "offense")) %>%
  filter(!is.na(offense)) %>% 
  pivot_wider(
    id_cols = geoid,
    names_from = offense,
    values_from = n,
    values_fill = 0) %>%
  left_join(census, .)

```
</div>

## Prepare data for plotting
<hr>

<div class = "now_you">
6\.	[[0.5]]{class="score"} Combine `crimes_by_census_tract`, `crimes`, and `census` into a single list object [[0.2]]{class="subscore"}. In doing so:

* [[0.2]]{class="subscore"} Within the list, assign the names `n_crimes` to `crimes_by_census_tract`, `crime_locations` to `crimes`, and `tracts` to `census`;
* [[0.1]]{class="subscore"} Globally assign the resultant object with the name `shapes_utm`.

```{r question_6}

shapes_utm <-
  list(
    "n_crimes" = crimes_by_census_tract,
    "crime_locations" = crimes,
    "tracts" = census)

```

7\.	[[1.5]]{class="score"} Using `purrr::map()` for iteration [[1.0]]{class="subscore"}, convert the CRS of the shapefiles contained in `shapes_utm` to EPSG 4326 [[0.4]]{class="subscore"} and assign the list to your global environment with the name `shapes_4326` [[0.1]]{class="subscore"}.

```{r question_7}

shapes_4326 <-
  shapes_utm %>%
  map(
    ~ st_transform(.x, crs = 4326))

```
</div>

## Generate maps of crime in DC
<hr>

<div class = "now_you">

8\.	[[1.5]]{class="score"} Using `shapes_utm`, generate a static choropleth tmap of census tracts [[0.5]]{class="subscore"} where the fill color is determined by the number of robberies committed [[1.0]]{class="subscore"}.

```{r question_8}

tm_shape(shapes_utm$n_crimes) +
tm_polygons(col = 'robbery')

```

9\. [[0.5]]{class="score"} Set the tmap mode to interactive viewing:

```{r question_9}

tmap_mode("view")

```

10\. [[1.0]]{class="score"} Using `shapes_4326`, generate an interactive tmap where:

* [[0.25]]{class="subscore"} The fill color of census tracts is determined by the number of robberies in a given tract;
* [[0.25]]{class="subscore"} Homicides are displayed as clusters of points;
* [[0.25]]{class="subscore"} OpenStreetMap and Esri.WorldImagery are provided as background
layers;
* [[0.25]]{class="subscore"} The layers are named "Robberies" and "Homicides".


```{r question_10}

tm_basemap(
  c("OpenStreetMap",
    "Esri.WorldImagery")) +
tm_shape(shapes_utm$n_crimes,
         name = "Robberies") +
tm_polygons(col = 'robbery') +
tm_shape(shapes_utm$n_crimes,
         name = "Homicides") +
tm_dots('homicide',
        clustering = TRUE) 
  
  

```
</div>
<hr>
**Extra credit!** [[0.25]]{class="score"} Modify Question 10 such that the polygons are semi-transparent (*Note: I have not taught transparency yet, but you can find information on how to do so with `?tm_polygons`*).

```{r extra_credit_1}


tm_basemap(
  c("OpenStreetMap",
    "Esri.WorldImagery")) +
tm_shape(shapes_utm$n_crimes,
         name = "Robberies") +
tm_polygons(col = 'robbery',
            alpha = .5) +
tm_shape(shapes_utm$n_crimes,
         name = "Homicides") +
tm_dots('homicide',
         clustering = TRUE) 
  

```

</div>

<hr>
 <script>
var acc = document.getElementsByClassName("accordion");
var i;

for (i = 0; i < acc.length; i++) {
  acc[i].addEventListener("click", function() {
    this.classList.toggle("active");
    var panel = this.nextElementSibling;
    if (panel.style.display === "block") {
      panel.style.display = "none";
    } else {
      panel.style.display = "block";
    }
  });
}
</script>