---
title: "Assignment, Week 5: Community science"
author: "Colleen Cronin"
date: "September 24, 2023"
output: html_document
---

<!-- Please knit this right away! The additional information for (html and css)
are provided to help you build on your current R Markdown toolbox. You will not
be responsible for learning these additional tools at this point. -->

<!-- The head tag, <head>, is a container for metadata in an html document. 
We can use it to define styles and do lots of other cool things things. -->
<head>
<!-- The link tag, <link>, creates a relationship between this file and
an external source. Here, I'm linking font-awesome so I can include a 
couple of icons that I like (user-secret and user-circle) -->
<link 
  rel="stylesheet" 
  href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">
<!-- Here I'm linking to a javascript file that will give us access to font awesome -->
  <script src="https://kit.fontawesome.com/03064bdd6c.js" crossorigin="anonymous"></script>
</head>
<!-- Note in the above that some html tags, like <head> need to be closed 
with </head>. Others, like <link> don't need a closing tag. -->

```{css, echo = FALSE}
.score {
  font-weight: bold;
  font-size: 16px;
  color: red;
  background-color: #ffff80;
}

.subscore {
  font-size: 14px;
  color: red;
}

/* Accordion style, the accordion is the section "Functions that you may use in this assignment" */
  
.accordion {
  background-color: #e6f0ff;
  color: #000000;
  cursor: pointer;
  padding: 18px;
  width: 100%;
  border-style: solid;
  border-width: 2px;
  border-color: #ffffff;
  text-align: left;
  outline: none;
  font-size: 15px;
  transition: 0.4s;
}

/* When you hover your mouse over the accordion button, it changes color. */

.active, .accordion:hover {
  background-color: #80b3ff; 
  color: #ffffff;
  font-weight: bold;
}

/* This is where the content of the accordion is placed. */

.panel {
  padding: 0 18px;
  display: none;
  background-color: #eeffff;
  overflow: hidden;
}

/* Here, I add a custom class for area surrounding the graded portions of this
assignment.*/

.now_you {
  border-left: 8px solid #292093;
  background-color: #e6ffff;
  padding: 16px;
  margin-top: 15px;
  margin-bottom: 15px;
}

/* This is where tips and tricks are formatted. */

.mysecret {
  border: 1px solid #999999;
  border-style: solid; 
  background-color: #efefef;
  padding: 20px; 
  border-radius: 25px;
  margin-top: 15px;
  margin-bottom: 15px;
}

h1, h2 {
  margin-bottom: 4px;
}

hr {
  margin-top: 0px;
}
```

<!-- Here is the setup for knitr. I had to include tidyverse so I could
include a tidyverse example with my Important warning statement below-->

```{r knitr_setup, include = FALSE}
options(knitr.table.format = "html")

knitr::opts_chunk$set(error = TRUE)
```

## Overview
<hr>

<a title = "iNat">
  <img  src = "https://boninabox.geobon.org/frontend/images/tool-images/inat1.jpg"
        alt="iNaturalist logo"
        style = "float: right; padding-left: 2em; padding-bottom: 1.5em; padding-right: 1em; width: 40%; height: auto;">
  </img>
</a>

In the last decade, technology-based community science programs, such as eBird and iNaturalist, have vastly expanded our understanding of species distributions. Data collected by program participants have altered range maps and revealed new insights on species' niches and the phenology of organisms (i.e., the timing of life history events). Concurrently, such programs also open opportunities to engage directly in the practice of collecting meaningful scientific data.

Despite the opportunities that such projects offer for scientists and the general public, the data often expose bias -- both in terms of sampling and among community science participants. For example, observations are often skewed towards areas with higher human population densities, public parks, and uncommon species. Concurrently, socioeconomic status is a key indicator of participation in these programs. This biases the distribution of observations but can also offer insights into how science can be made more accessible to the public.

In this exercise, you will explore data collected by iNaturalist participants in Washington, DC in May of 2022. You will use raster and point data to explore their observations.

<!-- A div is a container for html. I assign it to the class "my secret that
we created above -->
<div class = "mysecret">
<!-- The i tag is commonly used to create italic text, but it can also be 
used to create icons -->
<i class="fas fa-user-secret"></i> When I create an R Markdown file to communicate a coding process, I usually work in a .R script file and copy-and-paste the code into the R Markdown document at the end.
</div>

## Grading
<hr>

The points allotted for each question are provided in highlighted red bold text (e.g., <span class = "score">[1.0]</span>) within the question itself. When applicable, total points for a question may represent the sum of individually graded components, which are provided in red text (e.g., <span class = "subscore">[1.0]</span>).

Points may be deducted from each question's total: 

* [[50% per violation]]{class="subscore"} Use only functions listed in "Functions that you may use in this assignment";
* [[50%]]{class="subscore"} Include only assignments specified in the question;
* [[10%]]{class="subscore"} Ensure that all code is properly indented;
* [[10%]]{class="subscore"} Ensure that your code follows modern coding conventions (see *Best practices in modern R coding*);
* [[10%]]{class="subscore"} Ensure that all code chunks are formatted as `eval = TRUE` (but see below);
* [[10%]]{class="subscore"} Ensure that the class of your response is consistent with the wording of the question (e.g., if you are asked to return a "value" or "vector", ensure that the resultant object is not a data frame).

*Note: The maximum deduction is the total points value for a given question*

:::{class="mysecret"}
<i class="fas fa-user-secret fa-2x"></i> Pay careful attention to the format of your code -- not following the rules above will cost you lots of little points (and some big ones) that can really add up!
:::

In addition to points allotted per question, you must ensure that your R Markdown document **runs out-of-the-box** [[25% off of the total grade]]{class="subscore"} -- in other words, the document will knit without error. Some tips for doing so:

* Do not maintain your project folder in a location that is backed up by an online directory (e.g., Dropbox, iCloud).
* Ensure that all file paths are equivalent to those in this document (e.g., in `source()` or `read_csv()`).
* Do not use `setwd()` in your code (***Never** use `setwd()`!*)
* If one of your code chunks contains an error, include `eval = FALSE` in the options for that chunk. Otherwise, ensure all of your code chunks options are `eval = TRUE`.
* Prior to submission, clear all objects from your global environment and try to knit your document again.

:::{class="mysecret"}
<i class="fas fa-user-secret fa-2x"></i> Click the blue button below to view the functions that you may use in completing this problem set. Make sure that you know what each function does (use `?[function name]` if you do not). Do not use any functions outside of this list!
:::

<div>
<button class="accordion">Functions that you may use in this assignment</button>
<div class="panel">

In this assignment, you may use **only** the following R functions (*Note: If you are unclear on what a given function does, use `?` to view the help file!*):

* base::`=`
* base::`<-`
* base::`$`
* base::`~`
* base::`/`
* base::`c`
* base::`file.path`
* base::`library`
* base::`list.files`
* base::`mean`
* base::`names`
* base::`tolower`
* dplyr::`inner_join`
* dplyr::`mutate`
* dplyr::`n`
* dplyr::`pull`
* dplyr::`summarize`
* magrittr::`%>%`
* purrr::`map`
* rlang::`set_names`
* sf::`st_area`
* sf::`st_centroid`
* sf::`st_crs`
* sf::`st_difference`
* sf::`st_join`
* sf::`st_read`
* sf::`st_sf`
* sf::`st_transform`
* sf::`st_union`
* terra::`crop`
* terra::`extract`
* terra::`global`
* terra::`mask`
* terra::`rast`
* terra::`vect`
* tibble::`as_tibble`
* tmap::`+`
* tmap::`tmap_mode`
* tmap::`tm_dots`
* tmap::`tm_shape`
* tmap::`tm_polygons`
* tmap::`tm_raster`
* tmap::`tmap_mode`
* units::`set_units`

*Note: The packages dplyr, ggplot2, magrittr, readr, rlang, and tibble are all part of the tidyverse and are loaded with `library(tidyverse)`.*

</div>
</div>

## Getting started
<hr>

<div class = "now_you">

1\. [[0.5]]{class="score"} Save and knit this document:

* [[0.1]]{class="subscore"} Replace my name in the YAML header with yours
* [[0.1]]{class="subscore"} Add the current date in the YAML header
* [[0.3]]{class="subscore"} Save the *.rmd* file in the **output** folder of your project as (*but replace my name with yours*): `problem_set_5_Evans_Brian.rmd`

2\. [[0.5]]{class="score"} Load the `sf`, `tmap`, and `tidyverse` libraries.

```{r question_2, results = 'hide', message = FALSE, warning = FALSE}

library(sf)
library(tmap)
library(tidyverse)

```

3\. [[0.5]]{class="score"} Set the tmap mode to "view" for the entire document.

```{r question_3, message = FALSE, warning = FALSE}

tmap_mode('view')

```
</div>

## Data loading and pre-processing
<hr>


<div class = "now_you">
4\. [[1.0]]{class="score"} Modify the code below to such that it reads in the geojson shapefiles
`inat_2022_may`, `dc_surface_water`, and `dc_census` [[0.25]]{class="subscore"} and assign `shapes` to your global environment as a single list object. In doing so:

* [[0.25]]{class="subscore"} Assign the names `inat`, `water`, and `census` to the individual list items;
* [[0.25]]{class="subscore"} Convert the crs to EPSG 5070;
* [[0.25]]{class="subscore"} Set the fields of all files to lowercase.

```{r question_4, results = FALSE}
shapes <-
  c(
    inat = 'inat_2022_may.geojson',
    water = 'dc_surface_water.geojson',
    census = 'dc_census.geojson') %>% 
  map( 
        ~ file.path('data/raw/shapefiles', .x) %>% 
          st_read() %>% 
          set_names(
            names(.) %>% 
            tolower()) %>% 
          st_transform(crs = 5070))
```

*Note: The `income` variable is the median income per census tract.*

5\. [[1.0]]{class="score"} Using shapes$census:

* [[0.4]]{class="subscore"} Combine the census tracts into a single multipolygon;
* [[0.4]]{class="subscore"} Remove water from the resultant object;
* [[0.2]]{class="subscore"} Assign to your global environment with the name `dc_land`.

```{r question_5, results = FALSE}

dc_land_water <-
  shapes %>%
  map(~ .x %>%
        st_union() %>%
        st_sf())

dc_land <-
  dc_land_water$census %>%
  st_difference(dc_land_water$water)

rm(dc_land_water)

```

6\. [[1.0]]{class="score"} Using list.files to list the rasters in the folder 'data/rasters_problem_set_5' and a purrr::map function, modify the code below:

* [[0.25]]{class="subscore"} Read in the raster files;
* [[0.25]]{class="subscore"} Crop the rasters to shapes$census;
* [[0.25]]{class="subscore"} Convert the resultant object to a raster stack;
* [[0.25]]{class="subscore"} Mask the raster stack to shapes$census.

```{r question_6, results = FALSE}


rasters <-
  list.files("data/rasters_problem_set_5",
             full.names = TRUE) %>%
  set_names(c("canopy_cover",
              "impervious_surface")) %>%
  purrr::map(~ terra::rast(.x) %>%
               terra::crop(shapes$census) %>%
               terra::mask(shapes$census)) %>%
  terra::rast()



```
</div>

*Note: You may see a warning that the CRS of the two raster files are not equivalent -- don't worry, they are!*

## Exploring the data
<hr>

<div class = "now_you">
<i class="fas fa-user-circle"></i> 

7\. [[1.0]]{class="score"} Calculate the mean canopy cover of the **land area** in Washington DC. 

```{r question_7, message = FALSE, warning = FALSE}

dc_land %>%
  terra::vect() %>%
  terra::extract(
    rasters$canopy_cover,
    .,
    mean,
    na.rm = TRUE) %>%
  pull()



```

8\. [[1.5]]{class="score"} Generate a tmap [[0.5]]{class="subscore"} of mean canopy cover by census tract [[1.0]]{class="subscore"}.

```{r question_8, message = FALSE, warning = FALSE}

shapes$census %>%
  terra::vect() %>%
  terra::extract(
    rasters$canopy_cover,
    .,
    mean,
    na.rm = TRUE) %>%
  inner_join(
    shapes$census %>%
    mutate(ID = row_number()), .) %>%
  tm_shape() +
  tm_polygons(col = "canopy_cover",
              alpha = .6,
              palette = "Greens") 
```

9\. [[1.5]]{class="score"} Generate a choropleth tmap [[0.5]]{class="subscore"} that displays the density [[0.5]]{class="subscore"} of iNaturalist observations per census tract (number of observations per unit area; area unit = km^2) [[0.5]]{class="subscore"}.

```{r question_9, message = FALSE, warning = FALSE}

shapes$inat %>%
  st_join(shapes$census, .) %>%
  as_tibble() %>%
  summarize(n = n(),
            .by = geoid) %>%
  inner_join(shapes$census, .) %>%
  mutate(area =
           st_area(.) %>%
           units::set_units("km^2"),
           density = n / area) %>%
  tm_shape() +
  tm_polygons(
    title = 'Observations',
    style = 'kmeans',
    col = 'density',
    palette = 'viridis') 

```

<br>
10\. [[1.5]]{class="score"} Generate a tmap that displays (with layers ordered as follows):

* [[0.25]]{class="subscore"} Canopy cover;
* [[0.25]]{class="subscore"} Impervious surface cover;
* [[0.50]]{class="subscore"} The median income per census tract, displayed as polygons with the fill color determined by median income;
* [[0.50]]{class="subscore"} The number of iNaturalist observations per census tract displayed as centroids of tracts, with the point color determined by the total number of observations in each tract.

```{r question_10, message = FALSE, eval = FALSE, warning = FALSE}

inat_dense <-
  shapes$inat %>%
   st_join(shapes$census, .) %>%
   as_tibble() %>%
   summarize(n = n(),
            .by = geoid) %>%
   inner_join(shapes$census, .) %>% 
   mutate(area =
          st_area(.) %>%
          units::set_units("km^2"),
          density = n / area) 


# tm_basemap(
#   c('Esri.WorldTopoMap',
#     'OpenStreetMap',
#     'Esri.WorldImagery')) +

tm_shape(rasters$canopy_cover) +
tm_raster(alpha = .6,
    palette = "Greens") +

tm_shape(rasters$impervious_surface) +
tm_raster(alpha = .6,
    palette = "OrRd") +

tm_shape(shapes$census) +
tm_polygons(col = 'income',
    palette = 'viridis') +

tm_shape(inat_dense) +
tm_dots(col = 'density',
    palette = 'magma') 

```
</div>

**Extra credit** [[0.5]]{class="score"} : Use a custom color palette for each of the layers in Question 10!

<hr>

<script>
var acc = document.getElementsByClassName("accordion");
var i;

for (i = 0; i < acc.length; i++) {
  acc[i].addEventListener("click", function() {
    this.classList.toggle("active");
    var panel = this.nextElementSibling;
    if (panel.style.display === "block") {
      panel.style.display = "none";
    } else {
      panel.style.display = "block";
    }
  });
}
</script>